What We Eval in the Shadows

### 2021-04-09 paper/main.tex
tax

\medskip\noindent\EE{$min(e)=\c{F(X)}$} This form allows function calls whose
arguments may include variable references. Assignments are excluded from this
form. It occurs in 14\% of the sites in the corpus.

\medskip\noindent\EE{$min(e)=\c{F(F(X))}$} Nested function calls with variables.

\medskip\noindent\EE{$min(e)=\c{FUN}$} Function definition accounts for
  \packageFunctionDefinitionSitesPercent of \eval sites. In addition the
  functions inside other expressions account for
  \packageGeneralizedFunctionDefinitionSitesPercent of sites, and give an
  estimate of the use of higher-order functions.

\medskip\noindent\EE{$min(e)=\c{\{\,\}}$} Arbitrary-length statement blocks.
These are larger expressions with, on average, \packageMinimizedoperationsjRnd
instructions. The contents of the block are not inspected.

\medskip\noindent\EE{$min(e)=\c{model.frame}$} Call to the \c{model.frame}
function used to create statistical models. This subsume all other categories.

A given \eval call site can be passed different forms at different invocations.
However, \packageNbOneMinimizedPercent of sites only see a single form.

\todo{The following should be either removed or positioned properly}
We look at how \emph{consistent} the \c{expr} argument of \eval can be, \ie how
many different types of the resolved \c{expr} there are per call sites. Most of
the call sites, \ie \PercentMonomorphic, are \emph{consistent}, and this is
similar to javascript. However, a few ones are highly \emph{polymorphic} (10
different types). They are the pipe operators \c{\%>\%}, \c{\%<>\%} and
\c{\%\$\%} in package \emph{magrittr}. It is effectively used to compose
functions on their first argument, which can be of any type.

>>>>

\medskip\noindent\EE{$min(e)=\c{F(X)}$} This form allows for a simple function
call whose arguments may include variable references. Assignments are excluded
from this form. It occurs in 14\% of the sites in the corpus.

\medskip\noindent\EE{$min(e)=\c{F(F(X))}$} This form extends the previous one
with nested function calls. It is the second most common argument to \eval,
occurring in 20\% of the sites.

\medskip\noindent\EE{$min(e)=\c{FUN}$} This form captures expressions that are
function definitions, \eg~\c{function(x)x+1}. \c{FUN} occurs in
\packageFunctionDefinitionSitesPercent of sites. One data point we have not
recorded is the side effects that are performed when the generated functions are
eventually run. In addition to \c{FUN} there
\packageGeneralizedFunctionDefinitionSitesPercent of sites create functions
nested in other expressions, this gives an idea of the use of higher-order
functions.

\medskip\noindent\EE{$min(e)=\c{BLOCK}$} This form captures blocks of code with
multiple statements. These are larger expressions, we do not inspect contents of
the blocks. They occur in 4\% of the sites.

\medskip\noindent\EE{$min(e)=\c{model.frame}$} This is a very specific form
intended to capture the construction of calls to the \c{model.frame} function.
This function creates statistical models. This subsumes \c{F(F(X))}, \c{FUN} and
assignments. This particular form occurs in 7\% of the sites.

\paragraph{Consistency} It is interesting to consider how many different
forms any given site sees. The more forms, the harder it will be to characterize
the behavior of the program at that site. Luckily, \packageNbOneMinimizedPercent
of sites only see a single form. There are few sites that are highly polymorphic
(8 or more different forms), these include the pipe operator of the \c{magrittr}
package which is used to compose functions.

\paragraph{Discussion} The variety of uses of \eval is evidenced by the number
of different forms observed in CRAN. In comparison, JavaScript \eval usage was
simpler and more predictable as reported by \citet{oopsla12b}. Nevertheless,
simple forms dominate and there are many cases where \eval could be replaced
by less powerful constructs.
